---
title: "H5N1_SRs_Paper_t-tests"
output: html_document
date: "2024-09-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load libraries
```{r}
library(tidyverse)
library(openxlsx)
library(ggplot2)
library(wesanderson)
library(RColorBrewer)
library(rempsyc)
library(broom)
```

##Load raw data files
```{r}
raw_data_for_ttests <- '/Users/rgeiger/Dropbox/H5N1_SR_paper/SupplementaryTables_ttests.xlsx'
```

##Read sheets into tibbles
```{r}
read_sheet_data <- function(file,  sheetname) {
  sheet_data <- read.xlsx(file, sheet=sheetname, colNames=TRUE)
  return(sheet_data) }

figS1_tableS1 <- read_sheet_data(raw_data_for_ttests, 'Table_S1')
fig1B_tableS2 <- read_sheet_data(raw_data_for_ttests, 'Table_S2')
fig1C_tableS3 <- read_sheet_data(raw_data_for_ttests, 'Table_S3')
fig1D_tableS4 <- read_sheet_data(raw_data_for_ttests, 'Table_S4')
figS2B_tableS5 <- read_sheet_data(raw_data_for_ttests, 'Table_S5')
fig2C_tableS6 <- read_sheet_data(raw_data_for_ttests, 'Table_S6')
fig2D_tableS7 <- read_sheet_data(raw_data_for_ttests, 'Table_S7')
fig3A_tableS8 <- read_sheet_data(raw_data_for_ttests, 'Table_S8')
fig3B_tableS9 <- read_sheet_data(raw_data_for_ttests, 'Table_S9')
figS3A_tableS10 <- read_sheet_data(raw_data_for_ttests, 'Table_S10')
figS3B_tableS11 <- read_sheet_data(raw_data_for_ttests, 'Table_S11')
fig4A_tableS12 <- read_sheet_data(raw_data_for_ttests, 'Table_S12_analyzed')
fig4B_tableS13 <- read_sheet_data(raw_data_for_ttests, 'Table_S13_analyzed')
fig4C_tableS14 <- read_sheet_data(raw_data_for_ttests, 'Table_S14_analyzed')
fig4D_tableS15 <- read_sheet_data(raw_data_for_ttests, 'Table_S15_analyzed')
```

#Create Analyze Restriction Function
###Takes raw luciferase values, normalizes, calculates fold change to ev, and performs unpaired Welch's t-test to wtMxA
```{r}
analyze_restriction_data <- function(my_tbl) {
  ## get normalized values
  cols_to_use <- grep("firefly", colnames(my_tbl), value=TRUE)
  for (each_colname in cols_to_use) {
    renilla_colname <- gsub("firefly","renilla",each_colname)
    new_colname <- gsub("firefly-well","norm",each_colname)
    my_tbl[,new_colname] <- my_tbl[,each_colname] / my_tbl[,renilla_colname]
  }
  ## get EV mean norm
  ev_mean_norm <- my_tbl %>% 
    filter(variant == 'empty_vector') %>% 
    select(starts_with("norm")) %>% 
    as.numeric() %>% 
    na.omit() %>% 
    mean()
    
  ## calculate norm to EV
  norm_cols <- grep("norm", colnames(my_tbl), value = TRUE)
  for (each_colname in norm_cols) {
    norm_to_ev_colname <- gsub("norm", "norm_to_ev", each_colname)
    my_tbl[,norm_to_ev_colname] <- ev_mean_norm / my_tbl[,each_colname]
  }
  
  ## get just the norm_to_ev and variant cols
  my_tbl_selected <- my_tbl %>%
    select(variant, starts_with("norm_to_ev"))
  
  # Extract the control row based on the "variant" column
  if(!"wtMxA" %in% my_tbl_selected$variant) {
    stop("\n\nERROR - wtMxA not found in the variant column\n\n")
  }
  control_row <- my_tbl_selected %>%
    filter(variant == "wtMxA") %>%
    select(-variant) %>%
    slice(1) %>%  # Taking the first match in case there are duplicates
    as.numeric()
  
  # Function to perform t-test between control and a specific row
  t_test_with_control <- function(row) {
    t.test(control_row, row, paired = FALSE, var.equal = FALSE)
  }
  
  # Apply the function to each row excluding the control row
  ttest_results <- lapply(1:dim(my_tbl_selected)[1], function(row_number) {
    dat <- my_tbl_selected[row_number,] %>% 
      select(starts_with("norm_to_ev")) %>% 
      as.numeric()
    t_test_with_control(dat)
  })
  # return(ttest_results)
  ttest_results_by_variant <- sapply(names(ttest_results[[1]]), function(each_result)  {
    all_res <- sapply(ttest_results, "[[", each_result)
    if(class(all_res)[1] == "matrix") {
      all_res <- t(all_res)
      colnames(all_res) <- paste(each_result, 1:dim(all_res)[2], sep="_" )
      all_res <- as_tibble(all_res)
    } else {
      all_res <- tibble(myCol=all_res)
      colnames(all_res) <- each_result
    }
    return(all_res)
  })
  ttest_results_by_variant <- bind_cols(ttest_results_by_variant)
  
  my_tbl <- bind_cols(my_tbl, ttest_results_by_variant)%>% 
    relocate(p.value, .before = statistic)
  return(my_tbl)
}
```

##Results for single plate experiments
```{r}
fig1C_results <- analyze_restriction_data(fig1C_tableS3)

fig2C_results <- analyze_restriction_data(fig2C_tableS6)

fig2D_results <- analyze_restriction_data(fig2D_tableS7)

figS1_results <- analyze_restriction_data(figS1_tableS1)

figS2B_results <- analyze_restriction_data(figS2B_tableS5)

figS3A_results <- analyze_restriction_data(figS3A_tableS10)

figS3B_results <- analyze_restriction_data(figS3B_tableS11)
```


##Function for fold change to ev already calculated because of multiple plates
```{r}
##Note: both minireplicons have the 2X mxa listed as 200ng, but this is a typo. THOV got half the amount of DNA VN04 got
analyze_restriction_data_to2XMxA <- function(my_tbl) {
  ## get just the norm_to_ev and variant cols
  my_tbl_selected <- my_tbl %>%
    select(variant, starts_with("norm_to_ev"))
  
  # Extract the control row based on the "variant" column
  control_row <- my_tbl_selected %>%
    filter(variant == "wtMxA200") %>%
    select(-variant) %>%
    slice(1) %>%  # Taking the first match in case there are duplicates
    as.numeric()
  
  # Function to perform t-test between control and a specific row
  t_test_with_control <- function(row) {
    t.test(control_row, row, paired = FALSE, var.equal = FALSE)
  }
  
  # Apply the function to each row excluding the control row
  ttest_results <- lapply(1:dim(my_tbl_selected)[1], function(row_number) {
    dat <- my_tbl_selected[row_number,] %>% 
      select(starts_with("norm_to_ev")) %>% 
      as.numeric() %>% 
      na.omit()
    t_test_with_control(dat)
  })
  
  ttest_results_by_variant <- sapply(names(ttest_results[[1]]), function(each_result)  {
    all_res <- sapply(ttest_results, "[[", each_result)
    if(class(all_res)[1] == "matrix") {
      all_res <- t(all_res)
      colnames(all_res) <- paste(each_result, 1:dim(all_res)[2], sep="_" )
      all_res <- as_tibble(all_res)
    } else {
      all_res <- tibble(myCol=all_res)
      colnames(all_res) <- each_result
    }
    return(all_res)
  })
  
  ttest_results_by_variant <- bind_cols(ttest_results_by_variant)
  
  my_tbl <- bind_cols(my_tbl, ttest_results_by_variant) %>% 
    relocate(p.value, .before = statistic)
  return(my_tbl)
}

fig4A_results <- analyze_restriction_data_to2XMxA(fig4A_tableS12)

fig4B_results <- analyze_restriction_data_to2XMxA(fig4B_tableS13)
```

##For figure 2C t-tests to SWSTR
```{r}
analyze_restriction_data_toSWSTR <- function(my_tbl) {
  ## get normalized values
  cols_to_use <- grep("firefly", colnames(my_tbl), value=TRUE)
  for (each_colname in cols_to_use) {
    renilla_colname <- gsub("firefly","renilla",each_colname)
    new_colname <- gsub("firefly-well","norm",each_colname)
    my_tbl[,new_colname] <- my_tbl[,each_colname] / my_tbl[,renilla_colname]
  }
  ## get EV mean norm
  ev_mean_norm <- my_tbl %>% 
    filter(variant == 'empty_vector') %>% 
    select(starts_with("norm")) %>% 
    as.numeric() %>% 
    na.omit() %>% 
    mean()
    
  ## calculate norm to EV
  norm_cols <- grep("norm", colnames(my_tbl), value = TRUE)
  for (each_colname in norm_cols) {
    norm_to_ev_colname <- gsub("norm", "norm_to_ev", each_colname)
    my_tbl[,norm_to_ev_colname] <- ev_mean_norm / my_tbl[,each_colname]
  }
  
  ## get just the norm_to_ev and variant cols
  my_tbl_selected <- my_tbl %>%
    select(variant, starts_with("norm_to_ev"))
  
  # Extract the control row based on the "variant" column
  if(!"SWSTR" %in% my_tbl_selected$variant) {
    stop("\n\nERROR - SWSTR not found in the variant column\n\n")
  }
  control_row <- my_tbl_selected %>%
    filter(variant == "SWSTR") %>%
    select(-variant) %>%
    slice(1) %>%  # Taking the first match in case there are duplicates
    as.numeric()
  
  # Function to perform t-test between control and a specific row
  t_test_with_control <- function(row) {
    t.test(control_row, row, paired = FALSE, var.equal = FALSE)
  }
  
  # Apply the function to each row excluding the control row
  ttest_results <- lapply(1:dim(my_tbl_selected)[1], function(row_number) {
    dat <- my_tbl_selected[row_number,] %>% 
      select(starts_with("norm_to_ev")) %>% 
      as.numeric()
    t_test_with_control(dat)
  })
  # return(ttest_results)
  ttest_results_by_variant <- sapply(names(ttest_results[[1]]), function(each_result)  {
    all_res <- sapply(ttest_results, "[[", each_result)
    if(class(all_res)[1] == "matrix") {
      all_res <- t(all_res)
      colnames(all_res) <- paste(each_result, 1:dim(all_res)[2], sep="_" )
      all_res <- as_tibble(all_res)
    } else {
      all_res <- tibble(myCol=all_res)
      colnames(all_res) <- each_result
    }
    return(all_res)
  })
  ttest_results_by_variant <- bind_cols(ttest_results_by_variant)
  
  my_tbl <- bind_cols(my_tbl, ttest_results_by_variant)%>% 
    relocate(p.value, .before = statistic)
  return(my_tbl)
}

fig2C_results_toSWSTR <- analyze_restriction_data_toSWSTR(fig2C_tableS6)
```

##For figure 2C t-tests to SWSTR
```{r}
analyze_restriction_data_toNLS <- function(my_tbl) {
  ## get normalized values
  cols_to_use <- grep("firefly", colnames(my_tbl), value=TRUE)
  for (each_colname in cols_to_use) {
    renilla_colname <- gsub("firefly","renilla",each_colname)
    new_colname <- gsub("firefly-well","norm",each_colname)
    my_tbl[,new_colname] <- my_tbl[,each_colname] / my_tbl[,renilla_colname]
  }
  ## get EV mean norm
  ev_mean_norm <- my_tbl %>% 
    filter(variant == 'empty_vector') %>% 
    select(starts_with("norm")) %>% 
    as.numeric() %>% 
    na.omit() %>% 
    mean()
    
  ## calculate norm to EV
  norm_cols <- grep("norm", colnames(my_tbl), value = TRUE)
  for (each_colname in norm_cols) {
    norm_to_ev_colname <- gsub("norm", "norm_to_ev", each_colname)
    my_tbl[,norm_to_ev_colname] <- ev_mean_norm / my_tbl[,each_colname]
  }
  
  ## get just the norm_to_ev and variant cols
  my_tbl_selected <- my_tbl %>%
    select(variant, starts_with("norm_to_ev"))
  
  # Extract the control row based on the "variant" column
  if(!"NLS-wtMxA" %in% my_tbl_selected$variant) {
    stop("\n\nERROR - NLS-wtMxA not found in the variant column\n\n")
  }
  control_row <- my_tbl_selected %>%
    filter(variant == "NLS-wtMxA") %>%
    select(-variant) %>%
    slice(1) %>%  # Taking the first match in case there are duplicates
    as.numeric()
  
  # Function to perform t-test between control and a specific row
  t_test_with_control <- function(row) {
    t.test(control_row, row, paired = FALSE, var.equal = FALSE)
  }
  
  # Apply the function to each row excluding the control row
  ttest_results <- lapply(1:dim(my_tbl_selected)[1], function(row_number) {
    dat <- my_tbl_selected[row_number,] %>% 
      select(starts_with("norm_to_ev")) %>% 
      as.numeric()
    t_test_with_control(dat)
  })
  # return(ttest_results)
  ttest_results_by_variant <- sapply(names(ttest_results[[1]]), function(each_result)  {
    all_res <- sapply(ttest_results, "[[", each_result)
    if(class(all_res)[1] == "matrix") {
      all_res <- t(all_res)
      colnames(all_res) <- paste(each_result, 1:dim(all_res)[2], sep="_" )
      all_res <- as_tibble(all_res)
    } else {
      all_res <- tibble(myCol=all_res)
      colnames(all_res) <- each_result
    }
    return(all_res)
  })
  ttest_results_by_variant <- bind_cols(ttest_results_by_variant)
  
  my_tbl <- bind_cols(my_tbl, ttest_results_by_variant)%>% 
    relocate(p.value, .before = statistic)
  return(my_tbl)
}

figS2B_results_toNLS <- analyze_restriction_data_toNLS(figS2B_tableS5)
```

##For figure 4C-D t-tests to heterozygous
```{r}
analyze_restriction_data_toHet <- function(my_tbl) {
  ## get just the norm_to_ev and variant cols
  my_tbl_selected <- my_tbl %>%
    select(variant, starts_with("norm_to_ev"))
  
  # Extract the control row based on the "variant" column
  if(!"mwsrr50lydry50" %in% my_tbl_selected$variant) {
    stop("\n\nERROR - mwsrr50lydry50 not found in the variant column\n\n")
  }
  control_row <- my_tbl_selected %>%
    filter(variant == "mwsrr50lydry50") %>%
    select(-variant) %>%
    slice(1) %>%  # Taking the first match in case there are duplicates
    as.numeric()
  
  # Function to perform t-test between control and a specific row
  t_test_with_control <- function(row) {
    t.test(control_row, row, paired = FALSE, var.equal = FALSE)
  }
  
  # Apply the function to each row excluding the control row
  ttest_results <- lapply(1:dim(my_tbl_selected)[1], function(row_number) {
    dat <- my_tbl_selected[row_number,] %>% 
      select(starts_with("norm_to_ev")) %>% 
      as.numeric()
    t_test_with_control(dat)
  })
  # return(ttest_results)
  ttest_results_by_variant <- sapply(names(ttest_results[[1]]), function(each_result)  {
    all_res <- sapply(ttest_results, "[[", each_result)
    if(class(all_res)[1] == "matrix") {
      all_res <- t(all_res)
      colnames(all_res) <- paste(each_result, 1:dim(all_res)[2], sep="_" )
      all_res <- as_tibble(all_res)
    } else {
      all_res <- tibble(myCol=all_res)
      colnames(all_res) <- each_result
    }
    return(all_res)
  })
  ttest_results_by_variant <- bind_cols(ttest_results_by_variant)
  
  my_tbl <- bind_cols(my_tbl, ttest_results_by_variant)%>% 
    relocate(p.value, .before = statistic)
  return(my_tbl)
}

fig4C_results <- analyze_restriction_data_toHet(fig4C_tableS14)
fig4D_results <- analyze_restriction_data_toHet(fig4D_tableS15)
```

##Export results as new spreadsheet
```{r}
H5N1_SRs_paper_ttest_results <- createWorkbook()

addWorksheet(H5N1_SRs_paper_ttest_results, 'fig1C_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'fig2C_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'fig2C_results_toSWSTR')
addWorksheet(H5N1_SRs_paper_ttest_results, 'fig2D_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'figS1_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'figS2B_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'figS2B_results_toNLS')
addWorksheet(H5N1_SRs_paper_ttest_results, 'figS3A_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'figS3B_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'fig4A_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'fig4B_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'fig4C_results')
addWorksheet(H5N1_SRs_paper_ttest_results, 'fig4D_results')

writeData(H5N1_SRs_paper_ttest_results, 'fig1C_results', fig1C_results)
writeData(H5N1_SRs_paper_ttest_results, 'fig2C_results', fig2C_results)
writeData(H5N1_SRs_paper_ttest_results, 'fig2C_results_toSWSTR', fig2C_results_toSWSTR)
writeData(H5N1_SRs_paper_ttest_results, 'fig2D_results', fig2D_results)
writeData(H5N1_SRs_paper_ttest_results, 'figS1_results', figS1_results)
writeData(H5N1_SRs_paper_ttest_results, 'figS2B_results', figS2B_results)
writeData(H5N1_SRs_paper_ttest_results, 'figS2B_results_toNLS', figS2B_results_toNLS)
writeData(H5N1_SRs_paper_ttest_results, 'figS3A_results', figS3A_results)
writeData(H5N1_SRs_paper_ttest_results, 'figS3B_results', figS3B_results)
writeData(H5N1_SRs_paper_ttest_results, 'fig4A_results', fig4A_results)
writeData(H5N1_SRs_paper_ttest_results, 'fig4B_results', fig4B_results)
writeData(H5N1_SRs_paper_ttest_results, 'fig4C_results', fig4C_results)
writeData(H5N1_SRs_paper_ttest_results, 'fig4D_results', fig4D_results)

saveWorkbook(H5N1_SRs_paper_ttest_results, 'H5N1_SRs_paper_ttest_results.xlsx', overwrite = TRUE)
```